`r glue::glue('Doublet benchmarking results: {params$dataset}')`

### 2024-05-29 analyses/doublet-detection/template-notebooks/02_explore-benchmark-results.Rmd
left_join, density plots, and other styling

```{r functions}
plot_pca_d <- function(df, color_column, color_lab) {
  # Plot PC with a discrete color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, and `color_column`
  # The `color_column` argument should _not_ be provided as a string
  ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values = c("black", "lightblue")) + 
  # ensure doublets are on top for visibility
  geom_point(
    data = dplyr::filter(doublet_df, {{color_column}} == "doublet"), 
    color = "black"
  ) + 
  labs(color = color_lab) + 
  theme(legend.position = "bottom")
}

plot_pca_c <- function(df, color_column, pred_column, color_lab) {
  # Plot PC with a continuous color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, `color_column`, and `pred_column`
  # The `color_column` and `pred_column` arguments should _not_ be provided as strings
  ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  scale_color_viridis_c(direction = -1) + 
  # ensure doublets are on top for visibility
  geom_point(
    data = dplyr::filter(doublet_df, {{pred_column}} == "doublet"), 
    aes(color = {{color_column}})
  ) + 
  labs(color = color_lab) + 
  theme(legend.position = "bottom")
}

plot_doublet_scores <- function(df, score_col, pred_col, method_name){
  # df is expected to contain columns `ground_truth`, `score_col`, and `pred_col`
  # The column arguments should _not_ be provided as strings
  ggplot(df) +
    aes(x = ground_truth,
        y = {{score_col}}, 
        color = {{pred_col}}) + 
    geom_jitter(
      alpha = 0.2, 
      width = 0.1
    ) + 
    labs(
      x = "Ground truth call",
      y = glue::glue("{method_name} score"),
      color = glue::glue("{method_name} prediction")
    ) + 
  theme(legend.position = "bottom")
}

@@

doublet_df <- dplyr::inner_join(
  scdbl_df,
  scrub_df, 
  by = "barcodes"
) |>
  dplyr::inner_join(
    sce_df, 
    by = "barcodes"
  )
```

@@

```{r}
table(doublet_df$scdl_prediction)
```

@@

```{r fig.width = 8}
d_plot <- plot_pca_d(
  doublet_df, 
  scdbl_prediction,
  glue::glue("scDblFinder prediction")
)

c_plot <- plot_pca_c(
  doublet_df, 
  scdbl_score,
  scdbl_prediction,
  glue::glue("scDblFinder score")
)

@@

#### Doublet score distributions

```{r}
plot_doublet_scores(
  doublet_df,
  scdbl_score,
  scdbl_prediction,
  "scDblFinder prediction"
)
```

@@

@@

```{r fig.width = 8}
d_plot <- plot_pca_d(
  doublet_df, 
  scrublet_prediction,
  glue::glue("scrublet prediction")
)

c_plot <- plot_pca_c(
  doublet_df, 
  scrublet_score,
  scrublet_prediction,
  glue::glue("scrublet score")
)

@@

#### Doublet score distributions

```{r}
plot_doublet_scores(
  doublet_df,
  scrublet_score,
  scrublet_prediction,
  "scrublet prediction"
)
```

@@

> This method does not call droplets as singlet/doublet.
> Instead, performance is assessed below at several thresholds.

@@

##### Doublet score distributions

```{r}
plot_doublet_scores(
  doublet_df,
  cxds_score,
  cxds_prediction_0.5,
  "cxds prediction at 0.5"
)
```

@@

##### Doublet score distributions

```{r}
plot_doublet_scores(
  doublet_df,
  cxds_score,
  cxds_prediction_0.75,
  "cxds prediction at 0.75"
)
```

@@

##### Doublet score distributions

```{r}
plot_doublet_scores(
  doublet_df,
  cxds_score,
  cxds_prediction_0.9,
  "cxds prediction at 0.9"
)
```

>>>>

```{r functions}
plot_pca_d <- function(df, color_column, color_lab) {
  # Plot PC with a discrete color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, and `color_column`
  # The `color_column` argument should _not_ be provided as a string
  ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values = c("black", "lightblue")) + 
  # ensure doublets are on top for visibility
  geom_point(
    data = dplyr::filter(doublet_df, {{color_column}} == "doublet"), 
    color = "black"
  ) + 
  labs(color = color_lab) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

plot_pca_c <- function(df, color_column, pred_column, color_lab) {
  # Plot PC with a continuous color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, `color_column`, and `pred_column`
  # The `color_column` and `pred_column` arguments should _not_ be provided as strings
  ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  scale_color_viridis_c(direction = -1) + 
  # ensure doublets are on top for visibility
  geom_point(
    data = dplyr::filter(doublet_df, {{pred_column}} == "doublet"), 
    aes(color = {{color_column}})
  ) + 
  labs(color = color_lab) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

plot_jitter_scores <- function(df, score_col, pred_col, method_name){
  # Plot jitter plot of score distributions and predictions, colored by ground truth
  # df is expected to contain columns `ground_truth`, `score_col`, and `pred_col`
  # The column arguments should _not_ be provided as strings
  ggplot(df) +
    aes(x = ground_truth,
        y = {{score_col}}, 
        color = {{pred_col}}) + 
    geom_jitter(
      alpha = 0.2, 
      width = 0.1
    ) + 
    labs(
      x = "Ground truth call",
      y = glue::glue("{method_name} score"),
      color = glue::glue("{method_name} prediction")
    ) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

plot_density_scores <- function(df, score_col, method_name){
  # Plot density plot of score distributions, colored by ground truth
  # df is expected to contain columns `ground_truth` and `score_col`
  # The column arguments should _not_ be provided as strings
  ggplot(df) +
    aes(x = {{score_col}}, 
        fill = ground_truth) + 
    geom_density(alpha = 0.7) + 
    labs(
      x = glue::glue("{method_name} score"),
      fill = "Ground truth call"
    ) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

@@

# we can now remove the sce
rm(sce)

doublet_df <- dplyr::left_join(
  scdbl_df,
  scrub_df, 
  by = "barcodes"
) |>
  dplyr::left_join(
    sce_df, 
    by = "barcodes"
  )
```

@@

```{r}
table(doublet_df$scdbl_prediction)
```

#### Score distributions relative to ground truth

```{r}
plot_density_scores(
  doublet_df, 
  scdbl_score, 
  "scDblFinder"
)
```

@@

```{r fig.width = 8}
d_plot <- plot_pca_d(
  doublet_df, 
  scdbl_prediction,
  "scDblFinder prediction"
)

c_plot <- plot_pca_c(
  doublet_df, 
  scdbl_score,
  scdbl_prediction,
  "scDblFinder score"
)

@@

#### Jitter plot

```{r}
plot_jitter_scores(
  doublet_df,
  scdbl_score,
  scdbl_prediction,
  "scDblFinder prediction"
)
```

@@

#### Score distributions relative to ground truth

```{r}
plot_density_scores(
  doublet_df, 
  scrublet_score, 
  "scrublet"
)
```

@@

```{r fig.width = 8}
d_plot <- plot_pca_d(
  doublet_df, 
  scrublet_prediction,
  "scrublet prediction"
)

c_plot <- plot_pca_c(
  doublet_df, 
  scrublet_score,
  scrublet_prediction,
  "scrublet score"
)

@@

@@

Since this method does not call droplets as singlet/doublet, performance is assessed below at several thresholds: 0.5, 0.75, and 0.9.

#### Score distributions relative to ground truth

```{r}
plot_density_scores(
  doublet_df, 
  cxds_score, 
  "cxds"
)
```

@@

##### Jitter plot

```{r}
plot_jitter_scores(
  doublet_df,
  cxds_score,
  cxds_prediction_0.5,
  "cxds prediction at 0.5"
)
```

@@

##### Jitter plot

```{r}
plot_jitter_scores(
  doublet_df,
  cxds_score,
  cxds_prediction_0.75,
  "cxds prediction at 0.75"
)
```

@@

##### Jitter plot

```{r}
plot_jitter_scores(
  doublet_df,
  cxds_score,
  cxds_prediction_0.9,
  "cxds prediction at 0.9"
)
```
